---
title: "p8105_hw2_ak5098"
output: html_document
date: "2024-10-02"
---

## Homework 2

```{r}
library(tidyverse)
library(dplyr)
library(readr)
library(readxl)

subway = read_csv(file = "NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
                  col_types = cols(
                    Route8=col_character(),
                    Route9=col_character(),
                    Route10=col_character(),
                    Route11=col_character()))
```
```{r}
#Problem 1
subway = janitor::clean_names(subway)

subway_tidy = select(subway, line, station_name, station_latitude, station_longitude, starts_with("route"), entrance_type, entry, vending, ada)

subway_tidy %>% mutate(entry = case_match(entry,
  "YES" ~ TRUE, "NO" ~ FALSE, .default=NA))
```
## The dataset contains information about the station's name, exact location, route, entrance type and usability, vending and ADA accessibility. The dimensions are 1868x19. So far, I have only taken some cleaning steps. I have converted the names of the columns into more coding-friendly names, I have selected the columns I am interested in, and I have changed one column from a character column to logical. However, I have not yet removed missing values. I have also not changed other character columns to logical (when they could be converted). For these reasons, the data is not yet tidy. 


```{r}
distinct_station = distinct(subway, station_name)
nrows(distinct_station)

#356 distinct stations

distinct_station_ada = distinct(subway, station_name, entrance_type, ada, .keep_all = TRUE)
filter(distinct_station_ada, ada==TRUE) %>% print()

#137 ADA-compliant stations

no_vending = select(subway, station_name, entrance_type, vending, entry) %>% distinct(station_name, .keep_all=TRUE)  %>%  filter(vending=="NO", entry=="YES")
#9 station entrances/exits without vending
#5 station entrances/exits without vending allow entry
#proportion: 5/9 or 55.55% of stations entrances/exits without vending allow entry 
  
  
subway_pivot =  pivot_longer(
  subway,
  cols = starts_with("route"),
  names_to="route_number",
  values_to = "route_name",
  names_prefix = "route"
)

#Of the stations that serve A train, how many are ADA compliant?
subway_A = distinct(subway_pivot, station_name, route_name, .keep_all = TRUE) %>% filter(!is.na(route_name))
subway_A %>%  filter(route_name == "A") #56 unique stations that serve A train
subway_A %>%  filter(route_name == "A", ada==TRUE) #15 unique stations that serve A train and are ADA compliant
#15/56 or 26.79% of stations that serve the A train are ADA compliant
```

```{r}
#####PROBLEM 2
#tidying datasets
mrtrashwheel = read_excel("202409 Trash Wheel Collection Data.xlsx", 
                          sheet = "Mr. Trash Wheel", 
                          na = c("NA", "", "."))
mrtrashwheel_tidy = janitor::clean_names(mrtrashwheel)
mrtrashwheel_tidy = mrtrashwheel_tidy %>% mutate(sports_balls = as.integer(round(sports_balls, 0)))
mrtrashwheel_tidy = select(mrtrashwheel_tidy, -month, -year, -homes_powered, -x15, -x16) %>% drop_na() %>%  mutate(trashwheel="Mr Trash Wheel") 

professortrashwheel = read_excel("202409 Trash Wheel Collection Data.xlsx", 
                                 sheet = "Professor Trash Wheel",  
                                 na = c("NA", "", "."))
professortrashwheel_tidy = janitor::clean_names(professortrashwheel)
professortrashwheel_tidy = select(professortrashwheel_tidy, -month, -year, -homes_powered) %>% drop_na() %>% mutate(trashwheel="Professor Trash Wheel")

gwynnda = read_excel("202409 Trash Wheel Collection Data.xlsx", sheet = "Gwynnda Trash Wheel",  na = c("NA", "", ".")) 
gwynnda_tidy = janitor::clean_names(gwynnda)
gwynnda_tidy = select(gwynnda_tidy, -month, -year, -homes_powered) %>% drop_na() %>% mutate(trashwheel="Gwynnda")

#combining datasets

trashwheel_binded = bind_rows(mrtrashwheel_tidy, professortrashwheel_tidy, gwynnda_tidy, .id = "dumpster")
```

## Data Summary
This dataset contains information about each of the 117 dumpsters, including the date of the assessment, the weight and volume of the trash, and the contents of the trash. Trash is separated into categories of plastic bottles, polystyrene, cigarette butts, glass bottles, plastic bags, sports balls, and wrappers. 


```{r}
##PROBLEM 3
bakers = read_csv(file = "bakers.csv", na = c("NA", "", "."))
bakes = read_csv(file = "bakes.csv", na = c("NA", "", "."))
results = read_csv(file = "results.csv", na = c("NA", "", "."))
viewers = read_csv(file = "viewers.csv", na = c("NA", "", "."))

bakers_tidy = bakers %>% janitor::clean_names() %>% separate(baker_name, into = c("baker", "last_name"), sep = " ") %>% drop_na()
bakes_tidy = bakes %>%janitor::clean_names() %>% drop_na()
results_tidy = results %>%janitor::clean_names() %>% drop_na() %>% slice(-1:-2) %>%  rename(baker = x3, technical = x4, episode = x2, series = x1)

merged_bakers = inner_join(bakers_tidy, bakes_tidy,results_tidy, by = "baker")
merged_bakers = distinct(merged_bakers, baker, .keep_all = TRUE)


head(viewers)
viewers %>% janitor::clean_names()
mean(as.numeric(viewers$`Series 5`))


```
mean is 10.0393